# Introduction

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(latex2exp)
library(patchwork)
```

## Historical Introduction

### Exercises I 1.4 {#exr1.4}

**1.a** Modify the region in Figure I.3 by assuming that the ordinate at each $x$ is $2x^2$ instead of $x^2$. Draw the new figure. Check through the principal steps in the foregoing section and find what effect this has on the calculation of the area.

```{r echo=FALSE, warning=FALSE, fig.height=6, fig.width=7, fig.align='center'}

plot_integral_approx <- function(fun, 
                                 approx_type = c("below", "above"),
                                 y_labels = NULL,
                                 y_brake_init = 2,
                                 y_limit_inf = 0,
                                 fill_bars) {
    
    limit_inf <- 0
    limit_sup <- 1
    n_breaks  <- 6
    
    # Specify the type of approximation to apply
    approx_type <- switch(approx_type[[1]],
        "below" = 1,
        "above" = -1)
    
    # Specify bin width
    bin_width <- (limit_sup - limit_inf) / (n_breaks - 1)
    
    # Data
    tbl_data <- tibble::tibble(x = seq(from = {{limit_inf}}, 
                               to = {{limit_sup}}, 
                               length.out = {{n_breaks}})) %>% 
        # Define y-axis using the function fun
        dplyr::mutate(y = fun(x))
    
    # Plot 
    ggplot2::ggplot() +
        ggplot2::geom_col(data = tbl_data, aes(x, y), 
                 width = bin_width, 
                 # Approximation
                 ## from below:  1*bin_width
                 ## from above: -1*bin_width
                 position = position_nudge(x = (approx_type*bin_width) / 2),
                 color = "black",
                 fill  = fill_bars) + 
        # Plot the continuous function fun
        ggplot2::geom_function(fun = fun) +
        ggplot2::scale_x_continuous(breaks = seq(from = limit_inf, 
                                        to   = limit_sup,
                                        by   = bin_width), 
                           labels = c(latex2exp::TeX("$0$"),
                                      latex2exp::TeX("$\\frac{b}{n}$"),
                                      latex2exp::TeX("$\\frac{2b}{n}$"),
                                      latex2exp::TeX("$\\cdots$"),
                                      latex2exp::TeX("$\\frac{b(n-1)}{n}$"),
                                      latex2exp::TeX("$b")), 
                           limits = c(limit_inf, limit_sup), 
                           expand = c(0, 0)) + 
        ggplot2::scale_y_continuous(breaks = tbl_data$y[y_brake_init:length(tbl_data$y)],
                           labels = y_labels, 
                           expand = c(0, 0), 
                           limits = c(y_limit_inf, NA))  +
        ggplot2::theme(axis.text.y = element_text(hjust = 1),
              axis.text.x = element_text(vjust = 0.5), 
              axis.title  = element_blank(), 
              axis.line   = element_line(color = "black"),
              panel.background = element_blank())
}

parabola_2x2 <- function(x) {
    return(2*x^2)
}

approx_below_2x2 <- plot_integral_approx(fun = parabola_2x2, 
                     approx_type = "below", 
                     y_labels = c(expression(paste(2,bgroup("(", frac(b,n), ")")^2)),
                                  expression(paste(2,bgroup("(", frac(paste("2b"),n), ")")^2)),
                                  TeX("$\\cdots$"),
                                  expression(paste(2,bgroup("(", frac(paste("b(n-1)"),n), ")")^2)),
                                  TeX("$2b^2$")), 
                     y_brake_init = 2, 
                     fill_bars = "#a1d76a")

approx_above_2x2 <- plot_integral_approx(fun = parabola_2x2, 
                     approx_type = "above", 
                     y_labels = c(expression(paste(2,bgroup("(", frac(b,n), ")")^2)),
                                  expression(paste(2,bgroup("(", frac(paste("2b"),n), ")")^2)),
                                  TeX("$\\cdots$"),
                                  expression(paste(2,bgroup("(", frac(paste("b(n-1)"),n), ")")^2)),
                                  TeX("$2b^2$")), 
                     y_brake_init = 2, 
                     fill_bars = "#d8b365")

approx_below_2x2 + approx_above_2x2  
```

First we need to calculate the area of each rectangle:

$$\frac{b}{n} \mathbf{2}\left[\frac{kb}{n}\right]^2 = \mathbf{2}\frac{k^2b^3}{n^3}$$

Second we need to calculate $s_n$ 

$$\begin{split}
  s_n & = \sum_{k=1}^{n-1} \mathbf{2}\frac{k^2b^3}{n^3} \\
      & = \mathbf{2}\frac{b^3}{n^3} \sum_{k=1}^{n-1} k^2 \\
      & = \mathbf{2}\frac{b^3}{n^3} \left[\frac{n^3}{3} - \frac{n^2}{2} + \frac{n}{6} \right]
  \end{split}$$
  
Third we need to calculate $S_n$ 

$$\begin{split}
  S_n & = \sum_{k=1}^{n} \mathbf{2}\frac{k^2b^3}{n^3} \\
      & = \mathbf{2}\frac{b^3}{n^3} \sum_{k=1}^{n} k^2 \\
      & = \mathbf{2}\frac{b^3}{n^3} \left[\frac{n^3}{3} + \frac{n^2}{2} + \frac{n}{6} \right]
  \end{split}$$
  
Fourth we use the following inequalities where $n \geq 1$

$$\begin{split}
  \frac{n^3}{3} - \frac{n^2}{2} + \frac{n}{6} & < \frac{n^3}{3} & < \frac{n^3}{3} + \frac{n^2}{2} + \frac{n}{6} \\
  \mathbf{2}\frac{b^3}{n^3} \left[\frac{n^3}{3} - \frac{n^2}{2} + \frac{n}{6}\right] & < \mathbf{2}\frac{b^3}{3} & < \mathbf{2}\frac{b^3}{n^3} \left[\frac{n^3}{3} + \frac{n^2}{2} + \frac{n}{6}\right] \\
  s_n & < \mathbf{2}\frac{b^3}{3} & < S_n
  \end{split}$$

Fifth we assume that there is some $A$ such that $s_n < A < S_n$ and prove that $A = \mathbf{2}\frac{b^3}{3}$ using the following inequalities: $\frac{n^3}{3} - \frac{n^2}{2} + \frac{n}{6} < \frac{n^3}{3}$ and $\frac{n^3}{3} < \frac{n^3}{3} + \frac{n^2}{2} + \frac{n}{6}$

$$\begin{split}
  \frac{n^3}{3} - \frac{n^2}{2} + \frac{n}{6} < \frac{n^3}{3} & \iff \frac{n^3}{3} + \frac{n^2}{2} + \frac{n}{6} < \frac{n^3}{3} + n^2 \\
  & \iff \mathbf{2}\frac{b^3}{n^3}\left[\frac{n^3}{3} + \frac{n^2}{2} + \frac{n}{6}\right] < \mathbf{2}\frac{b^3}{n^3}\left[\frac{n^3}{3} + n^2\right] \\
  & \iff S_n < \mathbf{2}\left[\frac{b^3}{3} + \frac{b^3}{n}\right]
  \end{split}$$
  
  $$\begin{split}
  \frac{n^3}{3} < \frac{n^3}{3} + \frac{n^2}{2} + \frac{n}{6} & \iff \frac{n^3}{3} - n^2 < \frac{n^3}{3} - \frac{n^2}{2} + \frac{n}{6} \\
  & \iff \mathbf{2}\frac{b^3}{n^3}\left[\frac{n^3}{3} - n^2\right] < \mathbf{2}\frac{b^3}{n^3}\left[\frac{n^3}{3} - \frac{n^2}{2} + \frac{n}{6}\right] \\
  & \iff \mathbf{2}\left[\frac{b^3}{3} - \frac{b^3}{n}\right] < s_n \\
  \end{split}$$

Therefore we have also that $\mathbf{2}\left[\frac{b^3}{3} - \frac{b^3}{n}\right] < A < \mathbf{2}\left[\frac{b^3}{3} + \frac{b^3}{n}\right]$. If $A > \mathbf{2}\frac{b^3}{3}$ then $A - \mathbf{2}\frac{b^3}{3} > 0$ so using $A < \mathbf{2}\left[\frac{b^3}{3} + \frac{b^3}{n}\right]$ we can conclude that $n < \frac{\mathbf{2}b^3}{A - \mathbf{2}\frac{b^3}{3}}$ but $n \geq 1$ so there are some values of $n$ where $n < \frac{\mathbf{2}b^3}{A - \mathbf{2}\frac{b^3}{3}}$ is false. If $A < \mathbf{2}\frac{b^3}{3}$ then $0 < \mathbf{2}\frac{b^3}{3} - A$ so using $\mathbf{2}\left[\frac{b^3}{3} - \frac{b^3}{n}\right] < A$ we can also conclude that $n < \frac{\mathbf{2}b^3}{\mathbf{2}\frac{b^3}{3} - A}$ and again this is false. The only possibility is that $A = \mathbf{2}\frac{b^3}{3}$ so there is a unique value $A$ such that $s_n < A < S_n$.            

**1.b** Do the same if the ordinate at each $x$ is $3x^2$

```{r echo=FALSE, warning=FALSE, fig.height=6, fig.width=7, fig.align='center'}

parabola_3x2 <- function(x) {
    return(3*x^2)
}

approx_below_3x2 <- plot_integral_approx(fun = parabola_3x2, 
                                     approx_type = "below", 
                                     y_labels = c(expression(paste(3,bgroup("(", frac(b,n), ")")^2)),
                                  expression(paste(3,bgroup("(", frac(paste("2b"),n), ")")^2)),
                                  TeX("$\\cdots$"),
                                  expression(paste(3,bgroup("(", frac(paste("b(n-1)"),n), ")")^2)),
                                  TeX("$3b^2$")), 
                                     y_brake_init = 2, 
                                     fill_bars = "#a1d76a")

approx_above_3x2 <- plot_integral_approx(fun = parabola_3x2, 
                                     approx_type = "above", 
                                     y_labels = c(expression(paste(3,bgroup("(", frac(b,n), ")")^2)),
                                  expression(paste(3,bgroup("(", frac(paste("2b"),n), ")")^2)),
                                  TeX("$\\cdots$"),
                                  expression(paste(3,bgroup("(", frac(paste("b(n-1)"),n), ")")^2)),
                                  TeX("$3b^2$")), 
                                     y_brake_init = 2, 
                                     fill_bars = "#d8b365")

approx_below_3x2 + approx_above_3x2
```

The process is similar as in **1.a** where we only need to change $\mathbf{2}$ with $\mathbf{3}$.  

**1.c** Do the same if the ordinate at each $x$ is $\frac{1}{4}x^2$

```{r echo=FALSE, warning=FALSE, fig.height=6, fig.width=7, fig.align='center'}

parabola_1_4x2 <- function(x) {
    return((1/4)*x^2)
}

approx_below_1_4x2 <- plot_integral_approx(fun = parabola_1_4x2, 
                                     approx_type = "below", 
                                     y_labels = c(expression(paste(frac(paste("1"),paste("4")),bgroup("(", frac(b,n), ")")^2)),
                                  expression(paste(frac(paste("1"),paste("4")),bgroup("(", frac(paste("2b"),n), ")")^2)),
                                  TeX("$\\cdots$"),
                                  expression(paste(frac(paste("1"),paste("4")),bgroup("(", frac(paste("b(n-1)"),n), ")")^2)),
                                  TeX("$\\frac{1}{4}b^2$")), 
                                     y_brake_init = 2, 
                                     fill_bars = "#a1d76a")

approx_above_1_4x2 <- plot_integral_approx(fun = parabola_1_4x2, 
                                     approx_type = "above", 
                                     y_labels = c(expression(paste(frac(paste("1"),paste("4")),bgroup("(", frac(b,n), ")")^2)),
                                  expression(paste(frac(paste("1"),paste("4")),bgroup("(", frac(paste("2b"),n), ")")^2)),
                                  TeX("$\\cdots$"),
                                  expression(paste(frac(paste("1"),paste("4")),bgroup("(", frac(paste("b(n-1)"),n), ")")^2)),
                                  TeX("$\\frac{1}{4}b^2$")), 
                                     y_brake_init = 2, 
                                     fill_bars = "#d8b365")

approx_below_1_4x2 + approx_above_1_4x2
```

The process is similar as in **1.a** where you only need to change $\mathbf{2}$ with $\mathbf{\frac{1}{4}}$.

**1.d** Do the same if the ordinate at each $x$ is $2x^2 + 1$

```{r echo=FALSE, warning=FALSE, fig.height=6, fig.width=7, fig.align='center'}

parabola_2x2_1 <- function(x) {
    return(2*x^2 + 1)
}

approx_below_2x2_1 <- plot_integral_approx(fun = parabola_2x2_1, 
                                     approx_type = "below", 
                                     y_labels = c(TeX("$1$"),
                                                  expression(paste(2,bgroup("(", frac(b,n), ")")^2) + 1),
                                                  expression(paste(2,bgroup("(", frac(paste("2b"),n), ")")^2) + 1),
                                                  TeX("$\\cdots$"),
                                                  expression(paste(2,bgroup("(", frac(paste("b(n-1)"),n), ")")^2) + 1),
                                                  TeX("$2b^2 + 1$")), 
                                     y_brake_init = 1, 
                                     y_limit_inf = 0,   
                                     fill_bars = "#a1d76a")

approx_above_2x2_1 <- plot_integral_approx(fun = parabola_2x2_1, 
                                     approx_type = "above", 
                                     y_labels = c(TeX("$1$"),
                                                  expression(paste(2,bgroup("(", frac(b,n), ")")^2) + 1),
                                                  expression(paste(2,bgroup("(", frac(paste("2b"),n), ")")^2) + 1),
                                                  TeX("$\\cdots$"),
                                                  expression(paste(2,bgroup("(", frac(paste("b(n-1)"),n), ")")^2) + 1),
                                                  TeX("$2b^2 + 1$")), 
                                     y_brake_init = 1, 
                                     y_limit_inf = 0,  
                                     fill_bars = "#d8b365")

approx_below_2x2_1 + approx_above_2x2_1
```

First we need to calculate the area of each rectangle:

$$\frac{b}{n} \left[\mathbf{2}\left[\frac{kb}{n}\right]^2 + \mathbf{1}\right] = \mathbf{2}\frac{k^2b^3}{n^3} + \mathbf{1}\frac{b}{n}$$

Second we need to calculate $s_n$ 

$$\begin{split}
  s_n & = \sum_{k=1}^{n-1} \left[\mathbf{2}\frac{k^2b^3}{n^3} + \mathbf{1}\frac{b}{n}\right] \\
      & = \mathbf{2}\frac{b^3}{n^3} \sum_{k=1}^{n-1} k^2 + \mathbf{1}\frac{b(n-1)}{n} \\
      & = \mathbf{2}\frac{b^3}{n^3} \left[\frac{n^3}{3} - \frac{n^2}{2} + \frac{n}{6} \right] + \mathbf{1}\frac{b(n-1)}{n}
  \end{split}$$
  
Third we need to calculate $S_n$ 

$$\begin{split}
  S_n & = \sum_{k=1}^{n} \left[\mathbf{2}\frac{k^2b^3}{n^3} + \mathbf{1}\frac{b}{n}\right] \\
      & = \mathbf{2}\frac{b^3}{n^3} \sum_{k=1}^{n} k^2 + \mathbf{1}b \\
      & = \mathbf{2}\frac{b^3}{n^3} \left[\frac{n^3}{3} + \frac{n^2}{2} + \frac{n}{6} \right] + \mathbf{1}b
  \end{split}$$
  
Fourth we use the following inequalities where $n \geq 1$

$$\begin{split}
  \frac{n^3}{3} - \frac{n^2}{2} + \frac{n}{6} & < \frac{n^3}{3} < \frac{n^3}{3} + \frac{n^2}{2} + \frac{n}{6} \\
  \mathbf{2}\frac{b^3}{n^3} \left[\frac{n^3}{3} - \frac{n^2}{2} + \frac{n}{6}\right] + \mathbf{1}b & < \mathbf{2}\frac{b^3}{3} + \mathbf{1}b < \mathbf{2}\frac{b^3}{n^3} \left[\frac{n^3}{3} + \frac{n^2}{2} + \frac{n}{6}\right] + \mathbf{1}b \\
  \mathbf{2}\frac{b^3}{n^3} \left[\frac{n^3}{3} - \frac{n^2}{2} + \frac{n}{6}\right] + \mathbf{1}\frac{b(n-1)}{n} & < \mathbf{2}\frac{b^3}{3} + \mathbf{1}b < \mathbf{2}\frac{b^3}{n^3} \left[\frac{n^3}{3} + \frac{n^2}{2} + \frac{n}{6}\right] + \mathbf{1}b
  s_n & < \mathbf{2}\frac{b^3}{3} + \mathbf{1}b < S_n
  \end{split}$$

Fifth we assume that there is some $A$ such that $s_n < A < S_n$ and prove that $A = \mathbf{2}\frac{b^3}{3} + \mathbf{1}b$ using the following inequalities: $\frac{n^3}{3} - \frac{n^2}{2} + \frac{n}{6} < \frac{n^3}{3}$ and $\frac{n^3}{3} < \frac{n^3}{3} + \frac{n^2}{2} + \frac{n}{6}$

$$\begin{split}
  \frac{n^3}{3} - \frac{n^2}{2} + \frac{n}{6} < \frac{n^3}{3} & \iff \frac{n^3}{3} + \frac{n^2}{2} + \frac{n}{6} < \frac{n^3}{3} + n^2 \\
  & \iff \mathbf{2}\frac{b^3}{n^3}\left[\frac{n^3}{3} + \frac{n^2}{2} + \frac{n}{6}\right] + \mathbf{1}b < \mathbf{2}\frac{b^3}{n^3}\left[\frac{n^3}{3} + n^2\right] + \mathbf{1}b \\
  & \iff S_n < \mathbf{2}\left[\frac{b^3}{3} + \frac{b^3}{n}\right] + \mathbf{1}b
  \end{split}$$
  
  $$\begin{split}
  \frac{n^3}{3} < \frac{n^3}{3} + \frac{n^2}{2} + \frac{n}{6} & \iff \frac{n^3}{3} - n^2 < \frac{n^3}{3} - \frac{n^2}{2} + \frac{n}{6} \\
  & \iff \mathbf{2}\frac{b^3}{n^3}\left[\frac{n^3}{3} - n^2\right] + \mathbf{1}b < \mathbf{2}\frac{b^3}{n^3}\left[\frac{n^3}{3} - \frac{n^2}{2} + \frac{n}{6}\right] + \mathbf{1}b \\
  & \iff \mathbf{2}\left[\frac{b^3}{3} - \frac{b^3}{n}\right] + \mathbf{1}b < s_n \\
  \end{split}$$

Therefore we have also that $\mathbf{2}\left[\frac{b^3}{3} - \frac{b^3}{n}\right] + \mathbf{1}b < A < \mathbf{2}\left[\frac{b^3}{3} + \frac{b^3}{n}\right] + \mathbf{1}b$. If $A > \mathbf{2}\frac{b^3}{3} + \mathbf{1}b$ then $A - \mathbf{2}\frac{b^3}{3} - \mathbf{1}b > 0$ so using $A < \mathbf{2}\left[\frac{b^3}{3} + \frac{b^3}{n}\right] + \mathbf{1}b$ we can conclude that $n < \frac{\mathbf{2}b^3}{A - \mathbf{2}\frac{b^3}{3} - \mathbf{1}b}$ but $n \geq 1$ so there are some values of $n$ where $n < \frac{\mathbf{2}b^3}{A - \mathbf{2}\frac{b^3}{3} - \mathbf{1}b}$ is false. If $A < \mathbf{2}\frac{b^3}{3} + \mathbf{1}b$ then $0 < \mathbf{2}\frac{b^3}{3} + \mathbf{1}b - A$ so using $\mathbf{2}\left[\frac{b^3}{3} - \frac{b^3}{n}\right] + \mathbf{1}b < A$ we can also conclude that $n < \frac{\mathbf{2}b^3}{\mathbf{2}\frac{b^3}{3} + \mathbf{1}b - A}$ and again this is false. The only possibility is that $A = \mathbf{2}\frac{b^3}{3} + \mathbf{1}b$ so there is a unique value $A$ such that $s_n < A < S_n$.

**1.e** Do the same if the ordinate at each $x$ is $ax^2 + c$

In this part we are going to assume that $a > 0$ and $c > 0$. If $a = 0$ then it is not possible to calculate the area of the rectangles and if $a < 0$ we would need to define the concept of a negative area. Also if $c = 0$ it is a case we already cover and if $c < 0$ we would need again to define the concept of a negative area.    

```{r echo=FALSE, warning=FALSE, fig.height=6, fig.width=7, fig.align='center'}

parabola_ax2_c <- function(x) {
    return(2*x^2 + 1)
}

approx_below_ax2_c <- plot_integral_approx(fun = parabola_ax2_c, 
                                     approx_type = "below", 
                                     y_labels = c(TeX("$c$"),
                                                  expression(paste(a,bgroup("(", frac(b,n), ")")^2) + c),
                                                  expression(paste(a,bgroup("(", frac(paste("2b"),n), ")")^2) + c),
                                                  TeX("$\\cdots$"),
                                                  expression(paste(a,bgroup("(", frac(paste("b(n-1)"),n), ")")^2) + c),
                                                  TeX("$ab^2 + c$")), 
                                     y_brake_init = 1, 
                                     y_limit_inf = 0,   
                                     fill_bars = "#a1d76a")

approx_above_ax2_c <- plot_integral_approx(fun = parabola_ax2_c, 
                                     approx_type = "above", 
                                     y_labels = c(TeX("$c$"),
                                                  expression(paste(a,bgroup("(", frac(b,n), ")")^2) + c),
                                                  expression(paste(a,bgroup("(", frac(paste("2b"),n), ")")^2) + c),
                                                  TeX("$\\cdots$"),
                                                  expression(paste(a,bgroup("(", frac(paste("b(n-1)"),n), ")")^2) + c),
                                                  TeX("$ab^2 + c$")), 
                                     y_brake_init = 1, 
                                     y_limit_inf = 0,  
                                     fill_bars = "#d8b365")

approx_below_ax2_c + approx_above_ax2_c
```

The process is similar as in **1.d** where we only need to change $\mathbf{2}$ with $\mathbf{a}$ and $\mathbf{1}$ with $\mathbf{c}$.

**2** Modify the region in Figure I.3 by assuming that the ordinate at each $x$ is $x^3$ instead of $x^2$. Draw a new figure.

```{r echo=FALSE, warning=FALSE, fig.height=8, fig.width=7, fig.align='center'}

x3 <- function(x) {
    return(x^3)
}

approx_below_x3 <- plot_integral_approx(fun = x3, 
                                     approx_type = "below", 
                                     y_labels = c(expression(bgroup("(", frac(b,n), ")")^3),
                                                  expression(bgroup("(", frac(paste("2b"),n), ")")^3),
                                                  TeX("$\\cdots$"),
                                                  expression(bgroup("(", frac(paste("b(n-1)"),n), ")")^3),
                                                  TeX("$b^3$")), 
                                     y_brake_init = 2, 
                                     y_limit_inf = 0,   
                                     fill_bars = "#a1d76a")

approx_above_x3 <- plot_integral_approx(fun = x3, 
                                     approx_type = "above", 
                                     y_labels = c(expression(bgroup("(", frac(b,n), ")")^3),
                                                  expression(bgroup("(", frac(paste("2b"),n), ")")^3),
                                                  TeX("$\\cdots$"),
                                                  expression(bgroup("(", frac(paste("b(n-1)"),n), ")")^3),
                                                  TeX("$b^3$")), 
                                     y_brake_init = 2, 
                                     y_limit_inf = 0,  
                                     fill_bars = "#d8b365")

approx_below_x3 + approx_above_x3
```

**2a** Use a construction similar to that illustrated in Figure I.3 and show that the outer and inner sums $S_n$ and $s_n$ are given by

$$S_n = \frac{b^4}{n^4}\left[1^3 + 2^3 + \cdots + n^3\right] \; , \; s_n = \frac{b^4}{n^4}\left[1^3 + 2^3 + \cdots + (n-1)^3\right]$$

First we need to calculate the area of each rectangle:

$$\frac{b}{n} \left[\frac{kb}{n}\right]^3 = \frac{k^3b^4}{n^4}$$

Second we need to calculate $s_n$ 

$$\begin{split}
  s_n & = \sum_{k=1}^{n-1} \frac{k^3b^4}{n^4} \\
      & = \frac{b^4}{n^4} \sum_{k=1}^{n-1} k^3
  \end{split}$$
  
Third we need to calculate $S_n$ 

$$\begin{split}
  S_n & = \sum_{k=1}^{n} \frac{k^3b^4}{n^4} \\
      & = \frac{b^4}{n^4} \sum_{k=1}^{n} k^3
  \end{split}$$

**2b** Use the inequalities (which can be proved by mathematical induction; see Section I4.2)

$$1^3 + 2^3 + \cdots + (n-1)^3 < \frac{n^4}{4} < 1^3 + 2^3 + \cdots + n^3$$

to show that $s_n < \frac{b^4}{4} < S_n$ for every $n$, and prove that $\frac{b^4}{4}$ is the only number which lies between $s_n$ and $S_n$ for every $n$

First we use the following inequality where $n \geq 1$

$$\begin{split}
  1^3 + 2^3 + \cdots + (n-1)^3 & < \frac{n^4}{4} & < 1^3 + 2^3 + \cdots + n^3 \\
  \frac{b^4}{n^4} \sum_{k=1}^{n-1} k^3 & < \frac{b^4}{4} & < \frac{b^4}{n^4} \sum_{k=1}^{n} k^3 \\
  s_n & < \frac{b^4}{4} & < S_n
  \end{split}$$

Second we assume that there is some $A$ such that $s_n < A < S_n$ and prove that $A = \frac{b^4}{4}$ using the following inequalities: $s_n < \frac{b^4}{4}$ and $\frac{b^4}{4} < S_n$

If $A > \frac{b^4}{4}$ and $s_n < \frac{b^4}{4}$ then $S_n > \frac{b^4}{4}$ and $0 < \frac{b^4}{4} - s_n$. Therefore because $S_n - s_n > \frac{b^4}{4} - s_n$ this implies that $\frac{b^4}{\frac{b^4}{4} - s_n} > n$ but $n \geq 1$ so there are some values of $n$ where $\frac{b^4}{\frac{b^4}{4} - s_n} > n$ is false. 

If $A < \frac{b^4}{4}$ and $\frac{b^4}{4} < S_n$ then $\frac{b^4}{4} > s_n$ and $0 < S_n - \frac{b^4}{4}$. Therefore because $S_n - \frac{b^4}{4} < S_n - s_n$ this implies that $n < \frac{b^4}{S_n - \frac{b^4}{4}}$ but $n \geq 1$ so there are some values of $n$ where $n < \frac{b^4}{S_n - \frac{b^4}{4}}$ is false.

**2c** What numbers takes the place of $\frac{b^4}{4}$ if the ordinate at each $x$ is $ax^3 + c$?

The process to answer this question is similar to the answer in **2b** and **2a** 

First we need to calculate the area of each rectangle:

$$\frac{b}{n} \left[ a\left[\frac{kb}{n}\right]^{\mathbf{3}} + c\right] = a\frac{k^{\mathbf{3}}b^{\mathbf{4}}}{n^{\mathbf{4}}} + c \frac{b}{n}$$

Second we need to calculate $s_n$ 

$$\begin{split}
  s_n & = \sum_{k=1}^{n-1} \left[a\frac{k^{\mathbf{3}}b^{\mathbf{4}}}{n^{\mathbf{4}}} + c \frac{b}{n}\right] \\
      & = a\frac{b^{\mathbf{4}}}{n^{\mathbf{4}}} \sum_{k=1}^{n-1} k^{\mathbf{3}} + c\frac{b(n-1)}{n}
  \end{split}$$
  
Third we need to calculate $S_n$ 

$$\begin{split}
  S_n & = \sum_{k=1}^{n} \left[a\frac{k^{\mathbf{3}}b^{\mathbf{4}}}{n^{\mathbf{4}}} + c \frac{b}{n}\right] \\
      & = a\frac{b^{\mathbf{4}}}{n^{\mathbf{4}}} \sum_{k=1}^{n} k^{\mathbf{3}} + cb
  \end{split}$$

Fourth we use the following inequality where $n \geq 1$

$$\begin{split}
  1^{\mathbf{3}} + 2^{\mathbf{3}} + \cdots + (n-1)^{\mathbf{3}} & < \frac{n^{\mathbf{4}}}{{\mathbf{4}}} < 1^{\mathbf{3}} + 2^{\mathbf{3}} + \cdots + n^{\mathbf{3}} \\
  a\frac{b^{\mathbf{4}}}{n^{\mathbf{4}}} \sum_{k=1}^{n-1} k^{\mathbf{3}} + cb & < a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} + cb < a\frac{b^{\mathbf{4}}}{n^{\mathbf{4}}} \sum_{k=1}^{n} k^{\mathbf{3}} + cb \\
  a\frac{b^{\mathbf{4}}}{n^{\mathbf{4}}} \sum_{k=1}^{n-1} k^{\mathbf{3}} + c\frac{b(n-1)}{n} & < a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} + cb < a\frac{b^{\mathbf{4}}}{n^{\mathbf{4}}} \sum_{k=1}^{n} k^{\mathbf{3}} + cb \\
  s_n & < a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} + cb < S_n
  \end{split}$$

Sixth we assume that there is some $A$ such that $s_n < A < S_n$ and prove that $A = a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} + cb$ using the following inequalities: $s_n < a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} + cb$ and $a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} + cb < S_n$

If $A > a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} + cb$ and $s_n < a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} + cb$ then $S_n > a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} + cb$ and $0 < a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} + cb - s_n$. Therefore because $S_n - s_n > a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} + cb - s_n$ this implies that $\frac{ab^{\mathbf{4}} + cb}{a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} + cb - s_n} > n$ but $n \geq 1$ so there are some values of $n$ where $\frac{ab^{\mathbf{4}} + cb}{a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} + cb - s_n} > n$ is false. 

If $A < a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} + cb$ and $a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} + cb < S_n$ then $a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} + cb > s_n$ and $0 < S_n - a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} - cb$. Therefore because $S_n - a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} - cb < S_n - s_n$ this implies that $n < \frac{ab^4 + cb}{S_n - a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} - cb}$ but $n \geq 1$ so there are some values of $n$ where $n < \frac{ab^{\mathbf{4}} + cb}{S_n - a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} - cb}$ is false.

**3** The inequalities (I.5) and (I.12) are special cases of the more general inequalities

$$1^{\mathbf{m}} + 2^{\mathbf{m}} + \cdots + (n-1)^{\mathbf{m}} < \frac{n^{\mathbf{m + 1}}}{\mathbf{m + 1}} < 1^{\mathbf{m}} + 2^{\mathbf{m}} + \cdots + n^{\mathbf{m}}$$

that are valid for every integer $n \geq 1$ and every integer ${\mathbf{m}} \geq 1$. Asumme the validity of (I.13) and generalize the results of Exercise 2.

We can generalize this result to $ax^{\mathbf{m}} + c$ where the process is similar as in **2.c**. We only need to change $\mathbf{3}$ with $\mathbf{m}$ and $\mathbf{4}$ with $\mathbf{m + 1}$.

## Some Basic Concepts of the Theory of Sets

### Exercises I 2.5 {#exrI2.5}

**1** Use the roster notation to designate the following sets of real numbers.

$$A = \{x | x^2 - 1 = 0 \} = \{ -1, 1 \}$$

$$B = \{x | (x - 1)^2 = 0 \} = \{ 1 \}$$

$$C = \{x | x + 8 = 9 \} = \{ 1 \}$$

In this part $x^3 - 2x^2 + x = 2 \Longleftrightarrow (x^2 + 1)(x - 2) = 0$. In that sense there are also 2 complex roots $x = -i$ and $x = i$ but they are not real numbers (Remember that $S = \mathbb{R}$)

$$D = \{x | x^3 - 2x^2 + x = 2 \} = \{ 2 \}$$

$$E = \{x | (x + 8)^2 = 9^2 \} = \{ -17, 1 \}$$

$$F = \{x | (x^2 + 16x)^2 = 17^2 \} = \{ -17, -8 - \sqrt{47}, 1, \sqrt{47} - 8 \}$$

**2** For the sets in Exercise 1, note that $B \subseteq A$. List all the inclusion relations $\subseteq$ that hold among the sets $A$, $B$, $C$, $D$, $E$, $F$. 

$B \subseteq A$, $C \subseteq A$, $C \subseteq B$, $B \subseteq C$, $E \subseteq B$, $E \subseteq C$, $E \subseteq F$ 

**3** Let $A = \{ 1 \}$, $B = \{ 1, 2 \}$. Discuss the validity of the following statements (prove the ones that are true and explain why the others are not true).

**3a** $A \subset B$

It is true because $1 \in A$ and $1 \in B$ but $2 \notin A$ and $2 \in B$ so $A \subseteq B$ and $A \neq B$

**3b** $A \subseteq B$

It is true because $1 \in A$ and $1 \in B$ but $2 \notin A$ and $2 \in B$ so $A \subset B$ or $A = B$ is true

**3c** $A \in B$

It is false because ${1}$ is not an element of $B$. Remember that $1$ is not the same as $\{ 1 \}$

**3d** $1 \in A$

It is true because $1$ is an element of $A$. 

**3e** $1 \subseteq A$

It is false because the subsets of $A$ are $\emptyset$ and $\{ 1 \}$

**3f** $1 \subset B$

It is false because the proper subsets of $B$ are $\emptyset$, $\{ 1 \}$ and $\{ 2 \}$

**4** Solve Exercise 3 if $A = \{ 1 \}$ and $B = \{ \{ 1 \}, 1 \}$. 

**4a** $A \subset B$

It is true because $1 \in A$ and $1 \in B$ but $\{ 1 \} \notin A$ and $\{ 1 \} \in B$ so $A \subseteq B$ and $A \neq B$

**4b** $A \subseteq B$

It is true because $1 \in A$ and $1 \in B$ but $\{ 1 \} \notin A$ and $\{ 1 \} \in B$ so $A \subset B$ or $A = B$ is true

**4c** $A \in B$

It is true because $\{ 1 \} \in B$ and $A = \{ 1 \}$ so $A \in B$ 

**4d** $1 \in A$

It is true because $1$ is an element of $A$. 

**4e** $1 \subseteq A$

It is false because the subsets of $A$ are $\emptyset$ and $\{ 1 \}$

**4f** $1 \subset B$

It is false because the proper subsets of $B$ are $\emptyset$, $\{ \{1 \} \}$ and $\{ 1 \}$

**5** Given the set $S = \{ 1, 2, 3, 4 \}$. Display all subsets of $S$. There are 16 altogether, counting $\emptyset$ and $S$.

$\emptyset$, $\{ 1 \}$, $\{ 2 \}$, $\{ 3 \}$, $\{ 4 \}$, $\{ 1, 2 \}$, $\{ 1, 3 \}$, $\{ 1, 4 \}$, $\{ 2, 3 \}$, $\{ 2, 4 \}$, $\{ 3, 4 \}$, $\{ 1, 2, 3 \}$, $\{ 1, 2, 4 \}$, $\{ 1, 3, 4 \}$, $\{ 2, 3, 4 \}$, $\{ 1, 2, 3, 4 \}$

**6** Given the following four sets 

$$A = \{ 1, 2 \}, \;\; B = \{ \{ 1 \}, \{ 2 \} \}, \;\; C = \{ \{ 1 \}, \{ 1, 2 \} \}, \;\; D = \{ \{ 1 \}, \{ 2 \}, \{ 1, 2 \} \}$$  

discuss the validity of the following statements (prove the ones that are true and explain why the others are not true).

**6a** $A = B$

It is false because $1 \in A$ and $1 \notin B$

**6b** $A \subseteq B$

It is false because $1 \in A$ and $1 \notin B$

**6c** $A \subset C$

It is false because $1 \in A$ and $1 \notin C$

**6d** $A \in C$

It is true because $\{ 1, 2 \} \in C$ and $A = \{ 1, 2 \}$ so $A \in C$ 

**6e** $A \subset D$

It is false because $1 \in A$ and $1 \notin D$

**6f** $B \subset C$

It is false because $\{ 2 \} \in B$ and $\{ 2 \} \notin C$

**6g** $B \subset D$

It is true because $\{ 1 \} \in B$, $\{ 2 \} \in B$ and $\{ 1 \} \in D$, $\{ 2 \} \in D$ but $\{ 1, 2 \} \notin B$ and $\{ 1, 2 \} \in D$

**6h** $B \in D$

It is not true because the elements of $D$ are $\{ 1 \}$, $\{ 2 \}$ and $\{ 1, 2 \}$ so $\{ \{ 1 \}, \{ 2 \} \}$ is not an element of $D$

**6i** $A \in D$

It is true because $\{ 1, 2 \} \in D$ and $A = \{ 1, 2 \}$ so $A \in D$

**7** Prove the following properties of set equality.

**7a** $\{ a, a \} = \{ a \}$

$$\begin{split}
  a \in \{ a, a \} & \Longleftrightarrow a \in \{ a \}
  \end{split}$$

**7b** $\{ a, b \} = \{ b, a \}$

$$\begin{split}
  a \in \{ a, b \} \land b \in \{ a, b \} & \Longleftrightarrow b \in \{ a, b \} \land a \in \{ a, b \} \\
  & \Longleftrightarrow b \in \{ b, a \} \land a \in \{ b, a \}
  \end{split}$$

**7c** $\{ a \} = \{ b, c \}$ if and only if $a = b = c$

$$\begin{split}
  \{ a \} = \{ b, c \} & \Longleftrightarrow a \in \{ b, c \} \land (b \in \{ a \} \land c \in \{ a \}) \\
  & \Longleftrightarrow (a = b \lor a = c) \land (b = a \land c = a) \\
  & \Longleftrightarrow (a = b \lor a = c) \land (a = b = c) \\
  & \Longleftrightarrow a = b = c
  \end{split}$$

Prove the set relations in Exercises 8 through 19. (Sample proofs are given at the end of this section)

**8** *Commutative laws:* $A \cup B = B \cup A$, $A \cap B = B \cap A$

$$\begin{split}
  x \in A \cup B & \Longleftrightarrow x \in A \lor x \in B \\
  & \Longleftrightarrow x \in B \lor x \in A \\
  & \Longleftrightarrow \in B \cup A
  \end{split}$$

$$\begin{split}
  x \in A \cap B & \Longleftrightarrow x \in A \land x \in B \\
  & \Longleftrightarrow x \in B \land x \in A \\
  & \Longleftrightarrow \in B \cap A
  \end{split}$$

**9** *Associative laws:* $A \cup (B \cup C) = (A \cup B) \cup C$, $A \cap (B \cap C) = (A \cap B) \cap C$

$$\begin{split}
  x \in A \cup (B \cup C) & \Longleftrightarrow x \in A \lor x \in (B \cup C) \\
  & \Longleftrightarrow x \in A \lor (x \in B \lor x \in C) \\
  & \Longleftrightarrow (x \in A \lor x \in B) \lor x \in C \\
  & \Longleftrightarrow x \in A \cup B \lor x \in C \\
  & \Longleftrightarrow x \in (A \cup B) \cup C \\
  \end{split}$$

$$\begin{split}
  x \in A \cap (B \cap C) & \Longleftrightarrow x \in A \land x \in (B \cap C) \\
  & \Longleftrightarrow x \in A \land (x \in B \land x \in C) \\
  & \Longleftrightarrow (x \in A \land x \in B) \land x \in C \\
  & \Longleftrightarrow x \in A \cap B \land x \in C \\
  & \Longleftrightarrow x \in (A \cap B) \cap C \\
  \end{split}$$

**10** *Distributive laws:* $A \cap (B \cup C) = (A \cap B) \cup (A \cap C)$, $A \cup (B \cap C) = (A \cup B) \cap (A \cup C)$

$$\begin{split}
  x \in A \cap (B \cup C) & \Longleftrightarrow x \in A \land x \in (B \cup C) \\
  & \Longleftrightarrow x \in A \land (x \in B \lor x \in C) \\
  & \Longleftrightarrow (x \in A \land x \in B) \lor (x \in A \land x \in C) \\
  & \Longleftrightarrow x \in A \cap B \lor x \in A \cap C \\
  & \Longleftrightarrow x \in (A \cap B) \cup (A \cap C) \\
  \end{split}$$
  
$$\begin{split}
  x \in A \cup (B \cap C) & \Longleftrightarrow x \in A \lor x \in (B \cap C) \\
  & \Longleftrightarrow x \in A \lor (x \in B \land x \in C) \\
  & \Longleftrightarrow (x \in A \lor x \in B) \land (x \in A \lor x \in C) \\
  & \Longleftrightarrow x \in A \cup B \land x \in A \cup C \\
  & \Longleftrightarrow x \in (A \cup B) \cap (A \cup C) \\
  \end{split}$$

**11** $A \cup A = A$, $A \cap A = A$

$$\begin{split}
  x \in A \cup A & \Longleftrightarrow x \in A \lor x \in A \\
  & \Longleftrightarrow x \in A
  \end{split}$$
  
$$\begin{split}
  x \in A \cap A & \Longleftrightarrow x \in A \land x \in A \\
  & \Longleftrightarrow x \in A
  \end{split}$$
  
**12** $A \subseteq A \cup B$, $A \cap B \subseteq A$
 
$$\begin{split}
  x \in A & \Longrightarrow x \in A \lor x \in B \\
  & \Longrightarrow x \in A \cup B
  \end{split}$$
  
In this part we also have $A \cap B \subseteq B$ but because $A$ and $B$ are arbitrary sets the proof related to $A \cap B \subseteq A$ include this case. 
  
$$\begin{split}
  x \in A \cap B & \Longrightarrow x \in A \land x \in B \\
  & \Longrightarrow x \in A
  \end{split}$$
  
**13** $A \cup \emptyset = A$, $A \cap \emptyset = \emptyset$

In this part remember that $x \in \emptyset$ is false but $x \in A \lor x \in \emptyset$ is true if and only if $x \in A$  

$$\begin{split}
  x \in A \cup \emptyset & \Longleftrightarrow x \in A \lor x \in \emptyset \\
  & \Longleftrightarrow x \in A
  \end{split}$$

In this part remember that $\emptyset$ is always a subset of every set and that $\emptyset$ is a set containing no elements whatsoever. Taking into account these true statements we can prove that $A \cap \emptyset = \emptyset$.  

First, $A \cap \emptyset \supseteq \emptyset$ is true because $\emptyset$ is always a subset of every set.

Second, if $x \in A \cap \emptyset$ then $x \in A \land x \in \emptyset$ but $x \notin \emptyset$ so $x \in A \cap \emptyset$ is false. Therefore for any $x$ we have that $x \notin A \cap \emptyset$ must be true. Also let's assume that $A \cap \emptyset \nsubseteq \emptyset$ where this would mean that there is some $y$ such that $y \in A \cap \emptyset \land y \notin \emptyset$ but this is false because $A \cap \emptyset$ is a set containing no elements whatsoever. Therefore $A \cap \emptyset \nsubseteq \emptyset$ is false and $A \cap \emptyset \subseteq \emptyset$ must be true.

Third, because $A \cap \emptyset \supseteq \emptyset$ and $A \cap \emptyset \subseteq \emptyset$ are true we have that $A \cap \emptyset = \emptyset$. 

**14** $A \cup (A \cap B) = A$, $A \cap (A \cup B) = A$

First we prove $A \cup (A \cap B) = A$ in 2 parts:

$$\begin{split}
  x \in A \cup (A \cap B) & \Longleftrightarrow x \in A \lor x \in A \cap B \\
  & \Longleftrightarrow x \in A \lor (x \in A \land x \in B) \\
  & \Longleftrightarrow (x \in A \lor x \in A) \land (x \in A \lor x \in B) \\
  & \Longleftrightarrow x \in A \land x \in (A \cup B) \\
  & \Longrightarrow x \in A \\
  x \in A & \Longrightarrow x \in A \lor x \in A \cap B \\
  & \Longleftrightarrow x \in A \cup (A \cap B)
  \end{split}$$
  
$$\begin{split}
  x \in A & \Longrightarrow x \in A \lor x \in A \cap B \\
  & \Longleftrightarrow x \in A \cup (A \cap B)
  \end{split}$$

Second we prove $A \cap (A \cup B) = A$ using $A \cup (A \cap B) = A$:

$$\begin{split}
  A \cap (A \cup B) & = (A \cap A) \cup (A \cap B) \\
  & = A \cup (A \cap B) \\
  & = A
  \end{split}$$
  
**15** If $A \subseteq C$ and $B \subseteq C$, then $A \cup B \subseteq C$

$$\begin{split}
  (A \subseteq C \land B \subseteq C) \land x \in A \cup B & \Longleftrightarrow (A \subseteq C \land B \subseteq C) \land (x \in A \lor x \in B) \\
  & \Longrightarrow (A \subseteq C \land x \in A) \lor (B \subseteq C \land x \in B) \\
  & \Longrightarrow x \in C
  \end{split}$$

**16** If $C \subseteq A$ and $C \subseteq B$, then $C \subseteq A \cap B$

$$\begin{split}
  (C \subseteq A \land C \subseteq B) \land x \in C & \Longrightarrow (C \subseteq A \land x \in C) \land (C \subseteq B \land x \in C) \\
  & \Longrightarrow x \in A \land x \in B \\
  & \Longrightarrow x \in A \cap B
  \end{split}$$

**17a** If $A \subset B$ and $B \subset C$, prove that $A \subset C$

$$\begin{split}
  (A \subset B \land B \subset C) \land x \in A & \Longrightarrow (x \in B \land B \subset C) \land \exists y: y \notin A \land y \in B \\
  & \Longrightarrow x \in C \land \exists y: y \notin A \land y \in C
  \end{split}$$  

**17b** If $A \subseteq B$ and $B \subseteq C$, prove that $A \subseteq C$

$$\begin{split}
  (A \subseteq B \land B \subseteq C) \land x \in A & \Longrightarrow x \in B \land B \subseteq C \\
  & \Longrightarrow x \in C
  \end{split}$$

**17c** What can you conclude if $A \subset B$ and $B \subseteq C$?

First, if $B = C$ we can conclude that $A \subset C$.

Second, if $B \subset C$ using the proof in **17a** we can conclude again that $A \subset C$.

**17d** If $x \in A$ and $A \subseteq B$, is it necessarily true that $x \in B$?

Yes, because $A \subseteq B$ means that $\forall y: y \in A \Longrightarrow y \in B$ and if $x \in A$ is true also $x \in B$ must be true.

**17e** If $x \in A$ and $A \in B$ is it necessarily true that $x \in B$?

No, for example in **6** a counterexample was presented where $A = \{ 1, 2 \}$ and $C = \{ \{ 1\}, \{ 1, 2 \} \}$. In this case we have that $1 \in A$, $A \in C$ and $1 \notin C$.

**18** $A - (B \cap C) = (A - B) \cup (A - C)$

$$\begin{split}
  x \in A - (B \cap C) & \Longleftrightarrow x \in A \land x \notin B \cap C \\
  & \Longleftrightarrow x \in A \land (x \notin B \lor x \notin C) \\
  & \Longleftrightarrow (x \in A \land x \notin B) \lor (x \in A \land x \notin C) \\
  & \Longleftrightarrow x \in A - B \lor x \in A - C \\
  & \Longleftrightarrow x \in A - B \cup A - C
  \end{split}$$

**19** Let $\mathscr{F}$ be a class of sets. Then 

$$B - \bigcup_{A \in \mathscr{F}} A = \bigcap_{A \in \mathscr{F}} (B - A) \text{ and } B - \bigcap_{A \in \mathscr{F}} A = \bigcup_{A \in \mathscr{F}} (B - A)$$

$$\begin{split}
  x \in B - \bigcup_{A \in \mathscr{F}} A & \Longleftrightarrow x \in B \land x \notin \bigcup_{A \in \mathscr{F}} A \\
  & \Longleftrightarrow x \in B \land \forall A: A \in \mathscr{F} \land x \notin A \\
  & \Longleftrightarrow (\forall A: A \in \mathscr{F} \land x \in B) \land (\forall A: A \in \mathscr{F} \land x \notin A) \\
  & \Longleftrightarrow \forall A: A \in \mathscr{F} \land (x \in B \land x \notin A) \\
  & \Longleftrightarrow \forall A: A \in \mathscr{F} \land x \in B - A \\
  & \Longleftrightarrow x \in \bigcap_{A \in \mathscr{F}} B - A
  \end{split}$$

$$\begin{split}
  x \in B - \bigcap_{A \in \mathscr{F}} A & \Longleftrightarrow x \in B \land x \notin \bigcap_{A \in \mathscr{F}} A \\
  & \Longleftrightarrow x \in B \land \exists A: A \in \mathscr{F} \land x \notin A \\
  & \Longleftrightarrow (\exists A: A \in \mathscr{F} \land x \in B) \land (\exists A: A \in \mathscr{F} \land x \notin A) \\
  & \Longleftrightarrow \exists A: A \in \mathscr{F} \land (x \in B \land x \notin A) \\
  & \Longleftrightarrow \exists A: A \in \mathscr{F} \land x \in B - A \\
  & \Longleftrightarrow x \in \bigcup_{A \in \mathscr{F}} B - A
  \end{split}$$
  
**20a** Prove that one of the following two formulas is always right and the other one is sometimes wrong:

$$\text{(i) } A - (B - C) = (A - B) \cup C$$

$\text{(i)}$ is false because $A \cap C = C$ is not always true:

$$\begin{split}
  x \in A - (B - C) & \Longleftrightarrow x \in A \land x \notin B - C \\
  & \Longleftrightarrow x \in A \land (x \notin B \lor x \in C) \\
  & \Longleftrightarrow (x \in A \land x \notin B) \lor (x \in A \land x \in C) \\
  & \Longleftrightarrow x \in A - B \lor x \in A \cap C \\
  & \Longleftrightarrow x \in (A - B) \cup (A \cap C)
  \end{split}$$

$$\text{(ii) } A - (B \cup C) = (A - B) - C$$

$\text{(ii) }$ is true because:

$$\begin{split}
  x \in A - (B \cup C) & \Longleftrightarrow x \in A \land x \notin B \cup C \\
  & \Longleftrightarrow x \in A \land (x \notin B \land x \notin C) \\
  & \Longleftrightarrow (x \in A \land x \notin B) \land x \notin C \\
  & \Longleftrightarrow x \in A - B \land x \notin C \\
  & \Longleftrightarrow x \in (A - B) - C
  \end{split}$$

**20b** State an additional necessary and sufficient condition for the formula which is sometimes incorrect to be always right.

For $\text{(i)}$ to be true we need that $A \cap C = C$. Also because $A \cap C = C \Longleftrightarrow C \subseteq A$ another alternative is to point out that $C \subseteq A$ is a necessary and sufficient condition for $\text{(i)}$ to be true.   

## A Set of Axioms for the Real-Number System

### Exercises I 3.3 {#exrI3.3}

**1** Prove Theorems I.5 through I.15, using Axioms 1 through 6 and Theorems I.1 through I.4.

**Theorem I.5** $a(b - c) = ab - ac$

$$\begin{split}
  a(b - c) & = a(b  + (-c)) & \text{ Theorem I.3} \\
  & = ab  + a(-c) & \text{ Axiom 3} \\
  & = (ab + a(-c)) + 0 & \text{ Axiom 4} \\
  & = (ab + a(-c)) + (ac + (-ac)) & \text{ Theorem I.2} \\
  & = ab + (a(-c) + (ac + (-ac))) & \text{ Axiom 2} \\
  & = ab + ((a(-c) + ac) + (-ac)) & \text{ Axiom 2} \\
  & = ab + (a((-c) + c) + (-ac)) & \text{ Axiom 3} \\
  & = ab + (a( c + (-c)) + (-ac)) & \text{ Axiom 1} \\
  & = ab + (a0 + (-ac)) & \text{ Theorem I.2} \\
  & = (ab + a0) + (-ac) & \text{ Axiom 2} \\
  & = a(b + 0) + (-ac) & \text{ Axiom 3} \\
  & = ab + (-ac) & \text{ Axiom 4} \\
  & = ab -ac & \text{ Theorem I.3}
  \end{split}$$

**Theorem I.6** $0a = a0 = 0$

First 

$$0a = a0 \text{ Axiom 1}$$

Second

$$\begin{split}
  a0 & = a(0 + 0) & \text{ Axiom 4} \\
  & = a0 + a0 & \text{ Axiom 3}    
  \end{split}$$

Third 

$$\begin{split}
  a0 & = a0 + 0 \text{ Axiom 4}
  \end{split}$$

Fourth

$$\begin{split}
  a0 + a0 & = a0 + 0 \Longrightarrow a0 = 0  \text{ Theorem I.1}
  \end{split}$$

Therefore $0a = a0 = 0$

**Theorem I.7** CANCELLATION LAW OF MULTIPLICATION. If $ab = ac$ and $a \neq 0$, then $b = c$. (In particular, this shows that the number $1$ of Axiom 4 is unique)

First

$$\begin{split}
  a \neq 0 \Longrightarrow \exists y: y \in \mathbb{R} \land ay = 1 \text{ Axiom 6}
  \end{split}$$

Second

$$\begin{split}
  b & = 1b & \text{ Axiom 4} \\
  & = (ay)b \\
  & = (ya)b & \text{ Axiom 1} \\
  & = y(ab) & \text{ Axiom 2} \\
  & = y(ac) \\
  & = c & \text{ Axioms 2, 1, 4} \\
  \end{split}$$

**Theorem I.8** POSSIBILITY OF DIVISION. Given $a$ and $b$ with $a \neq 0$, there is exactly one $x$ such that $ax = b$. This $x$ is denoted by $b/a$ or $\frac{a}{b}$ and is called the quotient of $b$ and $a$. In particular, $1/a$ is also written $a^{-1}$ and is called the reciprocal of a.

First

$$\begin{split}
  a \neq 0 \Longrightarrow \exists y: y \in \mathbb{R} \land ay = 1 \text{ Theorem I.3}
  \end{split}$$

Second

$$\begin{split}
  b & = 1b & \text{ Axiom 4} \\
  & = (ay)b \\
  & = a(yb) & \text{ Axiom 2} \\
  & = ax & \text{ where } x = yb 
  \end{split}$$

Third, if $ax = b$ and $az = b$ then by **Theorem I.7** $x = z$ so $x$ is unique.

**Theorem I.9** If $a \neq 0$, then $b/a = ba^{-1}$

First

$$\begin{split}
  a \neq 0 \land 1 \in \mathbb{R} \Longrightarrow \exists! a^{-1}: a^{-1} \in \mathbb{R} \land aa^{-1} = 1 \text{ Theorem I.8}
  \end{split}$$

Second

$$\begin{split}
  a \neq 0 \land b \in \mathbb{R} \Longrightarrow \exists! b/a: b/a \in \mathbb{R} \land ab/a = b \text{ Theorem I.8}
  \end{split}$$

Third

$$\begin{split}
  b & = 1b & \text{ Axiom 4} \\
  & = (aa^{-1})b \\
  & = a(a^{-1}b) & \text{ Axiom 2} \\
  & = a(ba^{-1}) & \text{ Axiom 1} \\
  \end{split}$$

Fourth

$$\begin{split}
  ab/a = b \land b = a(ba^{-1}) \Longrightarrow b/a = ba^{-1} \text{ Theorem I.7}
  \end{split}$$

**Theorem I.10** If $a \neq 0$, then $(a^{-1})^{{-1}} = a$

First, if $a \neq 0$ we can use **Theorem I.8**

Second

$$\begin{split}
  1 & = a^{-1}(a^{-1})^{{-1}} \text{ Theorem I.8} \\
  \end{split}$$

Third

$$\begin{split}
  1 & = aa^{-1} & \text{ Theorem I.8} \\
  & = a^{-1}a & \text{ Axiom 1}
  \end{split}$$

Fourth

$$\begin{split}
  1 = a^{-1}(a^{-1})^{{-1}} \land 1 = a^{-1}a  \Longrightarrow (a^{-1})^{{-1}} = a \text{ Theorem I.7}
  \end{split}$$

**Theorem I.11** If $ab = 0$, then $a = 0$ or $b = 0$

To prove this statement we are going to use the method of *reductio ad absurdum* and the equivalence between the statements $A \Longrightarrow B$ and $\neg(A \land (\neg B))$. Please check out [@mendelson_number_2008, Chapter 1, p 7, Exercises 2 and 5]

First

$$\begin{split}
  (ab = 0 \Longrightarrow (a = 0 \lor b = 0)) & \Longleftrightarrow \neg(ab = 0 \land \neg(a = 0 \lor b = 0)) \\
  & \Longleftrightarrow \neg(ab = 0 \land (a \neq 0 \land b \neq 0))
  \end{split}$$

Second, we are going to assume $(ab = 0 \land (a \neq 0 \land b \neq 0)$, find a contradiction and conclude that $\neg(ab = 0 \land (a \neq 0 \land b \neq 0)$ is true where this means that the original statement is true: $ab = 0 \Longrightarrow (a = 0 \lor b = 0)$

Third

$$\begin{split}
  a0 = 0 \text{ Theorem I.6}
  \end{split}$$

Fourth

$$\begin{split}
  a0 = 0 \land ab = 0 \Longrightarrow b = 0 \text{ Theorem I.7}
  \end{split}$$

Fifth, there is a contradiction because we assume initially that $b \neq 0$ so $\neg(ab = 0 \land (a \neq 0 \land b \neq 0)$ is true by the method of *reductio ad absurdum*. Therefore the original statement is true: $ab = 0 \Longrightarrow (a = 0 \lor b = 0)$

**Theorem I.12** $(-a)b = -(ab)$ and $(-a)(-b) = ab$

In the case of $(-a)b = -(ab)$:

First

$$\begin{split}
  0 & = (-a)b + (-(-a)b) & \text{ Theorem I.8} \\
  & = (-a)b + (ab) & \text{ Theorem I.4} \\
  & = (ab) + (-a)b & \text{ Axiom 1}
  \end{split}$$

Second

$$\begin{split}
  0 & = (ab) + (-(ab)) \text{ Theorem I.8}
  \end{split}$$

Third

$$\begin{split}
  0 = (ab) + (-a)b \land 0 = (ab) + (-(ab)) \Longrightarrow (-a)b = (-(ab))  \text{ Theorem I.1}
  \end{split}$$

Fourth

$$\begin{split}
  -(ab) & = 0 - (ab) & \text{ Theorem I.2} \\
  & = 0 + (-(ab)) & \text{ Theorem I.3} \\
  & = (-(ab)) + 0 & \text{ Axiom 1} \\
  & = (-(ab)) & \text{ Axiom 4}
  \end{split}$$

Fifth

$$\begin{split}
  (-a)b & = (-(ab)) \\
  & = -(ab)
  \end{split}$$

In the case of $(-a)(-b) = ab$ we use the first part in **Theorem I.12**:

$$\begin{split}
  (-a)(-b) & = -(a(-b)) \\
  & = -((-b)a) & \text{ Axiom 1} \\
  & = (-(-b))a \\
  & = -(-b)a & \text{ Theorem I.2, Theorem I.3, Axiom 4} \\
  & = ba & \text{ Theorem I.4} \\
  & = ab & \text{ Axiom 1}
  \end{split}$$

**Theorem I.13** $(a / b) + (c / d) = (ad + bc) / (bd)$ if $b \neq 0$ and $d \neq 0$

First we prove that $bd(b^{-1}d^{-1}) = 1$

$$\begin{split}
  bd(b^{-1}d^{-1}) & = bb^{-1}(dd^{-1}) \text{ Axiom 2, Axiom 1} \\
  & = 1\cdot1 \text{ Theorem I.8} \\
  & = 1 \text{ Axiom 4}
  \end{split}$$

Second we prove that $(b^{-1}d^{-1}) = (bd)^{-1}$ taking into account that $bd(bd)^{-1} = 1$ by **Theorem I.8** 

$$\begin{split}
  bd(b^{-1}d^{-1}) = 1 \land bd(bd)^{-1} = 1 \Longrightarrow (b^{-1}d^{-1}) = (bd)^{-1} \text{ Theorem I.11, Theorem I.7}  
  \end{split}$$


Third

$$\begin{split}
  (a / b) + (c / d) & = (ab^{-1}) + (cd^{-1}) & \text{ Theorem I.9} \\
  & = 1(ab^{-1}) + 1(cd^{-1}) & \text{ Axiom 4} \\
  & = dd^{-1}(ab^{-1}) + bb^{-1}(cd^{-1}) & \text{ Theorem I.8, Theorem I.9} \\
  & = (b^{-1}d^{-1})ad + (b^{-1}d^{-1})bc & \text{ Axiom 2, Axiom 1} \\
  & = (ad + bc)(b^{-1}d^{-1}) & \text{ Axiom 3, Axiom 1} \\
  & = (ad + bc)(bd)^{-1} \\
  & = (ad + bc) / (bd) & \text{  Theorem I.9}
  \end{split}$$
  
**Theorem I.14** $(a / b)(c / d) = (ac) / (bd)$ if $b \neq 0$ and $d \neq 0$

First, check out the proof of **Theorem I.13** where we prove that $(b^{-1}d^{-1}) = (bd)^{-1}$

Second

$$\begin{split}
  (a / b)(c / d) & = (ab^{-1})(cd^{-1}) & \text{ Theorem I.9} \\
  & = (ac)(b^{-1}d^{-1}) & \text{ Axiom 2, Axiom 1} \\
  & = (ac)(bd)^{-1} & \text{ Axiom 2, Axiom 1} \\
  & = (ac) / (bd) & \text{ Theorem I.9}
  \end{split}$$

**Theorem I.15** $(a / b) / (c / d) = (ad) / (bc)$ if $b \neq 0$, $c \neq 0$ and $d \neq 0$ 

First, check out the proof of **Theorem I.13** where we prove that $(b^{-1}d^{-1}) = (bd)^{-1}$. Using this result we can also prove that $(b^{-1}(d^{-1})^{-1}) = (bd^{-1})^{-1}$ where $b$ and $d$ are any real numbers different from $0$

Second

$$\begin{split}
  (a / b) / (c / d) = (ad) / (bc) & = (ab^{-1}) / (cd^{-1}) & \text{ Theorem I.9} \\
  & = (ab^{-1})(cd^{-1})^{-1} & \text{ Theorem I.9} \\
  & = (ab^{-1})(c^{-1}(d^{-1})^{-1}) \\
  & = (ab^{-1})(c^{-1}d) & \text{ Theorem I.10} \\
  & = (ad)(b^{-1}c^{-1}) & \text{ Axiom 2, Axiom 1} \\
  & = (ad)(bc)^{-1} \\
  & = (ad) / (bc) & \text{ Theorem I.9}
  \end{split}$$

In Exercises 2 through 10, prove the given statements or establish the given questions. You may use Axioms 1 through 6 and Theorems I.1 through I.15.

**2** $-0 = 0$

$$\begin{split}
  -0 \land 0 & \Longleftrightarrow 0 + (-0) = 0 \land 0 + 0 = 0 & \text{ Axiom 4, Theorem I.2} \\
  & \Longleftrightarrow (-0) = 0 & \text{ Theorem I.1} \\
  & \Longleftrightarrow -0 = 0 & \text{ Theorem I.3, Theorem I.2, Axiom 4}
  \end{split}$$

**3** $1^{-1} = 1$

$$\begin{split}
  1^{-1} \land 1 & \Longleftrightarrow 1\cdot 1^{-1} = 1 \land 1 \cdot 1 = 1 & \text{ Axiom 4, Axiom 6, Theorem I.8} \\
  & \Longleftrightarrow 1^{-1} = 1 & \text{ Theorem I.7}
  \end{split}$$

**4** Zero has no reciprocal

To prove this statement we are going to use the method of *reductio ad absurdum*. Please check out [@mendelson_number_2008, Chapter 1, p 7, Exercises 2 and 5]

First, let's assume that $x$ is the reciprocal of $0$. Then by **Theorem 1.8** $0 \cdot x = 1$.

Second by **Theorem 1.6** $0 \cdot x = 0$ but we are assuming that $0 \cdot x = 1$ so $0 = 1$.

Third, there is a contradiction because by **Axiom 4** $0$ and $1$ are distinct.

Fourth $0 \cdot x = 1$ is false and $0 \cdot x \neq 1$ is true so zero has no reciprocal.

**5** $-(a + b) = -a - b$

$$\begin{split}
  -(a + b) & = -1(a + b) & \text{ Axiom 4} \\
  & = -1a + (-1b) & \text{ Axiom 3} \\
  & = -a + (-b) & \text{ Axiom 4} \\
  & = -a - b & \text{ Theorem I.3} \\
  \end{split}$$

**6** $-(a - b) = -a + b$

$$\begin{split}
  -(a - b) & = -1(a - b) & \text{ Axiom 4} \\
  & = -1a + (-1)(-b) & \text{ Axiom 3} \\
  & = -1a + 1b & \text{ Theorem I.12} \\
  & = -a + b & \text{ Axiom 4}
  \end{split}$$

**7** $(a - b) + (b - c) = a + c$

$$\begin{split}
  (a - b) + (b - c) & = (a + (-b)) + (b + (-c)) & \text{ Theorem I.3} \\
  & = (a + ((-b) + b)) + (-c) & \text{ Axiom 2} \\
  & = (a + 0) + (-c) & \text{ Axiom 1, Axiom 4} \\
  & = a - c & \text{ Axiom 4, Theorem I.3} \\
  \end{split}$$

**8** If $a \neq 0$ and $b \neq 0$, then $(ab)^{-1} = (a^{-1}b^{-1})$

Check out the proof of **Theorem I.13** where we prove that $(b^{-1}d^{-1}) = (bd)^{-1}$

**9** $-(a / b) = (-a / b) = a / (-b)$ if $b \neq 0$

First we prove that $-1 \neq 0$. If $-1 = 0$ then $1 + (-1) = 0$ and $0 + 0 = 0$ will imply that $1 = 0$ which is not true by **Axiom 4**. Therefore by *reductio ad absurdum* (please check out [@mendelson_number_2008, Chapter 1, p 7, Exercises 2 and 5]) $-1 \neq 0$ must be true. 

Second we prove that $-b^{-1} = (-b)^{-1}$. Because $(-b)(-b)^1 = 1$ by **Theorem I.8**, $(-b)(-b^{-1}) = bb^{-1} = 1$ by **Theorem I.12, Theorem I.8** and $-b = -1b \neq 0$ by **Theorem I.11** we can conclude that $-b^{-1} = (-b)^{-1}$ using **Theorem I.7**. 

$$\begin{split}
  -(a / b) & = -(ab^{-1}) & \text{ Theorem I.9} \\
  & = (-ab^{-1}) & \text{ Theorem I.3, Axiom 4} \\
  & = (-a / b) & \text{ Theorem I.9} \\
  & = (-ab^{-1}) & \text{ Theorem I.9} \\
  & = (-b^{-1}a) & \text{ Axiom 1} \\
  & = -(b^{-1}a) & \text{ Theorem I.3, Axiom 4} \\
  & = (-b^{-1})a & \text{ Theorem I.12} \\
  & = (-b)^{-1}a \\
  & = a(-b)^{-1} & \text{ Axiom 1} \\
  & = a / (-b) & \text{ Theorem I.9}
  \end{split}$$
  
  **10** $(a / b) - (c / d) = (ad - bc) / (bd)$ if $b \neq 0$ and $d \neq 0$
  
  In this part remember exercise **8**

$$\begin{split}
  (a / b) - (c / d) & = (ab^{-1}) - (cd^{-1}) & \text{ Theorem I.9} \\
  & = 1(ab^{-1}) - 1(cd^{-1}) & \text{ Axiom 4} \\
  & = dd^{-1}(ab^{-1}) - bb^{-1}(cd^{-1}) & \text{ Theorem I.8, Theorem I.9} \\
  & = (b^{-1}d^{-1})ad - (b^{-1}d^{-1})bc & \text{ Axiom 2, Axiom 1} \\
  & = (ad - bc)(b^{-1}d^{-1}) & \text{ Axiom 3, Theorem I.5, Axiom 1} \\
  & = (ad - bc)(bd)^{-1} \\
  & = (ad - bc) / (bd) & \text{  Theorem I.9}
  \end{split}$$

### Exercises I 3.5 {#exrI3.5}

### Exercises I 3.12 {#exrI3.12}

## Mathematical Induction, Summation Notation, and Related Topics

### Exercises I 4.4 {#exrI4.4}

### Exercises I 4.7 {#exrI4.7}

### Exercises I 4.9 {#exrI4.9}
