# Introduction

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(latex2exp)
library(patchwork)
```

## Historical Introduction

### Exercises I 1.4 {#exrI1.4}

**1.a** Modify the region in Figure I.3 by assuming that the ordinate at each $x$ is $2x^2$ instead of $x^2$. Draw the new figure. Check through the principal steps in the foregoing section and find what effect this has on the calculation of the area.

```{r echo=FALSE, warning=FALSE, fig.height=6, fig.width=7, fig.align='center'}

plot_integral_approx <- function(fun, 
                                 approx_type = c("below", "above"),
                                 y_labels = NULL,
                                 y_brake_init = 2,
                                 y_limit_inf = 0,
                                 fill_bars) {
    
    limit_inf <- 0
    limit_sup <- 1
    n_breaks  <- 6
    
    # Specify the type of approximation to apply
    approx_type <- switch(approx_type[[1]],
        "below" = 1,
        "above" = -1)
    
    # Specify bin width
    bin_width <- (limit_sup - limit_inf) / (n_breaks - 1)
    
    # Data
    tbl_data <- tibble::tibble(x = seq(from = {{limit_inf}}, 
                               to = {{limit_sup}}, 
                               length.out = {{n_breaks}})) %>% 
        # Define y-axis using the function fun
        dplyr::mutate(y = fun(x))
    
    # Plot 
    ggplot2::ggplot() +
        ggplot2::geom_col(data = tbl_data, aes(x, y), 
                 width = bin_width, 
                 # Approximation
                 ## from below:  1*bin_width
                 ## from above: -1*bin_width
                 position = position_nudge(x = (approx_type*bin_width) / 2),
                 color = "black",
                 fill  = fill_bars) + 
        # Plot the continuous function fun
        ggplot2::geom_function(fun = fun) +
        ggplot2::scale_x_continuous(breaks = seq(from = limit_inf, 
                                        to   = limit_sup,
                                        by   = bin_width), 
                           labels = c(latex2exp::TeX("$0$"),
                                      latex2exp::TeX("$\\frac{b}{n}$"),
                                      latex2exp::TeX("$\\frac{2b}{n}$"),
                                      latex2exp::TeX("$\\cdots$"),
                                      latex2exp::TeX("$\\frac{b(n-1)}{n}$"),
                                      latex2exp::TeX("$b")), 
                           limits = c(limit_inf, limit_sup), 
                           expand = c(0, 0)) + 
        ggplot2::scale_y_continuous(breaks = tbl_data$y[y_brake_init:length(tbl_data$y)],
                           labels = y_labels, 
                           expand = c(0, 0), 
                           limits = c(y_limit_inf, NA))  +
        ggplot2::theme(axis.text.y = element_text(hjust = 1),
              axis.text.x = element_text(vjust = 0.5), 
              axis.title  = element_blank(), 
              axis.line   = element_line(color = "black", 
                                         arrow = grid::arrow(ends = "last",
                                                       length = unit(0.05, "inches"))),
              panel.background = element_blank())
}

parabola_2x2 <- function(x) {
    return(2*x^2)
}

approx_below_2x2 <- plot_integral_approx(fun = parabola_2x2, 
                     approx_type = "below", 
                     y_labels = c(TeX("$2\\left(\\frac{b}{n}\\right)^2$"),
                                  TeX("$2\\left(\\frac{2b}{n}\\right)^2$"),
                                  TeX("$\\cdots$"),
                                  TeX("$2\\left(\\frac{b(n-1)}{n}\\right)^2$"),
                                  TeX("$2b^2$")), 
                     y_brake_init = 2, 
                     fill_bars = "#a1d76a")

approx_above_2x2 <- plot_integral_approx(fun = parabola_2x2, 
                     approx_type = "above", 
                     y_labels = c(TeX("$2\\left(\\frac{b}{n}\\right)^2$"),
                                  TeX("$2\\left(\\frac{2b}{n}\\right)^2$"),
                                  TeX("$\\cdots$"),
                                  TeX("$2\\left(\\frac{b(n-1)}{n}\\right)^2$"),
                                  TeX("$2b^2$")), 
                     y_brake_init = 2, 
                     fill_bars = "#d8b365")

approx_below_2x2 + approx_above_2x2  
```

First we need to calculate the area of each rectangle:

$$\frac{b}{n} \mathbf{2}\left[\frac{kb}{n}\right]^2 = \mathbf{2}\frac{k^2b^3}{n^3}$$

Second we need to calculate $s_n$ 

$$\begin{split}
  s_n & = \sum_{k=1}^{n-1} \mathbf{2}\frac{k^2b^3}{n^3} \\
      & = \mathbf{2}\frac{b^3}{n^3} \sum_{k=1}^{n-1} k^2 \\
      & = \mathbf{2}\frac{b^3}{n^3} \left[\frac{n^3}{3} - \frac{n^2}{2} + \frac{n}{6} \right]
  \end{split}$$
  
Third we need to calculate $S_n$ 

$$\begin{split}
  S_n & = \sum_{k=1}^{n} \mathbf{2}\frac{k^2b^3}{n^3} \\
      & = \mathbf{2}\frac{b^3}{n^3} \sum_{k=1}^{n} k^2 \\
      & = \mathbf{2}\frac{b^3}{n^3} \left[\frac{n^3}{3} + \frac{n^2}{2} + \frac{n}{6} \right]
  \end{split}$$
  
Fourth we use the following inequalities where $n \geq 1$

$$\begin{split}
  \frac{n^3}{3} - \frac{n^2}{2} + \frac{n}{6} & < \frac{n^3}{3} & < \frac{n^3}{3} + \frac{n^2}{2} + \frac{n}{6} \\
  \mathbf{2}\frac{b^3}{n^3} \left[\frac{n^3}{3} - \frac{n^2}{2} + \frac{n}{6}\right] & < \mathbf{2}\frac{b^3}{3} & < \mathbf{2}\frac{b^3}{n^3} \left[\frac{n^3}{3} + \frac{n^2}{2} + \frac{n}{6}\right] \\
  s_n & < \mathbf{2}\frac{b^3}{3} & < S_n
  \end{split}$$

Fifth we assume that there is some $A$ such that $s_n < A < S_n$ and prove that $A = \mathbf{2}\frac{b^3}{3}$ using the following inequalities: $\frac{n^3}{3} - \frac{n^2}{2} + \frac{n}{6} < \frac{n^3}{3}$ and $\frac{n^3}{3} < \frac{n^3}{3} + \frac{n^2}{2} + \frac{n}{6}$

$$\begin{split}
  \frac{n^3}{3} - \frac{n^2}{2} + \frac{n}{6} < \frac{n^3}{3} & \iff \frac{n^3}{3} + \frac{n^2}{2} + \frac{n}{6} < \frac{n^3}{3} + n^2 \\
  & \iff \mathbf{2}\frac{b^3}{n^3}\left[\frac{n^3}{3} + \frac{n^2}{2} + \frac{n}{6}\right] < \mathbf{2}\frac{b^3}{n^3}\left[\frac{n^3}{3} + n^2\right] \\
  & \iff S_n < \mathbf{2}\left[\frac{b^3}{3} + \frac{b^3}{n}\right]
  \end{split}$$
  
  $$\begin{split}
  \frac{n^3}{3} < \frac{n^3}{3} + \frac{n^2}{2} + \frac{n}{6} & \iff \frac{n^3}{3} - n^2 < \frac{n^3}{3} - \frac{n^2}{2} + \frac{n}{6} \\
  & \iff \mathbf{2}\frac{b^3}{n^3}\left[\frac{n^3}{3} - n^2\right] < \mathbf{2}\frac{b^3}{n^3}\left[\frac{n^3}{3} - \frac{n^2}{2} + \frac{n}{6}\right] \\
  & \iff \mathbf{2}\left[\frac{b^3}{3} - \frac{b^3}{n}\right] < s_n \\
  \end{split}$$

Therefore we have also that $\mathbf{2}\left[\frac{b^3}{3} - \frac{b^3}{n}\right] < A < \mathbf{2}\left[\frac{b^3}{3} + \frac{b^3}{n}\right]$. If $A > \mathbf{2}\frac{b^3}{3}$ then $A - \mathbf{2}\frac{b^3}{3} > 0$ so using $A < \mathbf{2}\left[\frac{b^3}{3} + \frac{b^3}{n}\right]$ we can conclude that $n < \frac{\mathbf{2}b^3}{A - \mathbf{2}\frac{b^3}{3}}$ but $n \geq 1$ so there are some values of $n$ where $n < \frac{\mathbf{2}b^3}{A - \mathbf{2}\frac{b^3}{3}}$ is false. If $A < \mathbf{2}\frac{b^3}{3}$ then $0 < \mathbf{2}\frac{b^3}{3} - A$ so using $\mathbf{2}\left[\frac{b^3}{3} - \frac{b^3}{n}\right] < A$ we can also conclude that $n < \frac{\mathbf{2}b^3}{\mathbf{2}\frac{b^3}{3} - A}$ and again this is false. The only possibility is that $A = \mathbf{2}\frac{b^3}{3}$ so there is a unique value $A$ such that $s_n < A < S_n$.            

**1.b** Do the same if the ordinate at each $x$ is $3x^2$

```{r echo=FALSE, warning=FALSE, fig.height=6, fig.width=7, fig.align='center'}

parabola_3x2 <- function(x) {
    return(3*x^2)
}

approx_below_3x2 <- plot_integral_approx(fun = parabola_3x2, 
                                     approx_type = "below", 
                                     y_labels = c(TeX("$3\\left(\\frac{b}{n}\\right)^2$"),
                                                  TeX("$3\\left(\\frac{2b}{n}\\right)^2$"),
                                                  TeX("$\\cdots$"),
                                                  TeX("$3\\left(\\frac{b(n-1)}{n}\\right)^2$"),
                                                  TeX("$3b^2$")), 
                                     y_brake_init = 2, 
                                     fill_bars = "#a1d76a")

approx_above_3x2 <- plot_integral_approx(fun = parabola_3x2, 
                                     approx_type = "above", 
                                     y_labels = c(TeX("$3\\left(\\frac{b}{n}\\right)^2$"),
                                                  TeX("$3\\left(\\frac{2b}{n}\\right)^2$"),
                                                  TeX("$\\cdots$"),
                                                  TeX("$3\\left(\\frac{b(n-1)}{n}\\right)^2$"),
                                                  TeX("$3b^2$")), 
                                     y_brake_init = 2, 
                                     fill_bars = "#d8b365")

approx_below_3x2 + approx_above_3x2
```

The process is similar as in **1.a** where we only need to change $\mathbf{2}$ with $\mathbf{3}$.  

**1.c** Do the same if the ordinate at each $x$ is $\frac{1}{4}x^2$

```{r echo=FALSE, warning=FALSE, fig.height=6, fig.width=7, fig.align='center'}

parabola_1_4x2 <- function(x) {
    return((1/4)*x^2)
}

approx_below_1_4x2 <- plot_integral_approx(fun = parabola_1_4x2, 
                                     approx_type = "below", 
                                     y_labels = c(TeX("$\\frac{1}{4}\\left(\\frac{b}{n}\\right)^2$"),
                                                  TeX("$\\frac{1}{4}\\left(\\frac{2b}{n}\\right)^2$"),
                                                  TeX("$\\cdots$"),
                                                  TeX("$\\frac{1}{4}\\left(\\frac{b(n-1)}{n}\\right)^2$"),
                                                  TeX("$\\frac{1}{4}b^2$")), 
                                     y_brake_init = 2, 
                                     fill_bars = "#a1d76a")

approx_above_1_4x2 <- plot_integral_approx(fun = parabola_1_4x2, 
                                     approx_type = "above", 
                                     y_labels = c(TeX("$\\frac{1}{4}\\left(\\frac{b}{n}\\right)^2$"),
                                                  TeX("$\\frac{1}{4}\\left(\\frac{2b}{n}\\right)^2$"),
                                                  TeX("$\\cdots$"),
                                                  TeX("$\\frac{1}{4}\\left(\\frac{b(n-1)}{n}\\right)^2$"),
                                                  TeX("$\\frac{1}{4}b^2$")), 
                                     y_brake_init = 2, 
                                     fill_bars = "#d8b365")

approx_below_1_4x2 + approx_above_1_4x2
```

The process is similar as in **1.a** where you only need to change $\mathbf{2}$ with $\mathbf{\frac{1}{4}}$.

**1.d** Do the same if the ordinate at each $x$ is $2x^2 + 1$

```{r echo=FALSE, warning=FALSE, fig.height=6, fig.width=7, fig.align='center'}

parabola_2x2_1 <- function(x) {
    return(2*x^2 + 1)
}

approx_below_2x2_1 <- plot_integral_approx(fun = parabola_2x2_1, 
                                     approx_type = "below", 
                                     y_labels = c(TeX("$1$"),
                                                  TeX("$2\\left(\\frac{b}{n}\\right)^2 + 1$"),
                                                  TeX("$2{1}{4}\\left(\\frac{2b}{n}\\right)^2 + 1$"),
                                                  TeX("$\\cdots$"),
                                                  TeX("$2\\left(\\frac{b(n-1)}{n}\\right)^2 + 1$"),
                                                  TeX("$2b^2 + 1$")), 
                                     y_brake_init = 1, 
                                     y_limit_inf = 0,   
                                     fill_bars = "#a1d76a")

approx_above_2x2_1 <- plot_integral_approx(fun = parabola_2x2_1, 
                                     approx_type = "above", 
                                     y_labels = c(TeX("$1$"),
                                                  TeX("$2\\left(\\frac{b}{n}\\right)^2 + 1$"),
                                                  TeX("$2{1}{4}\\left(\\frac{2b}{n}\\right)^2 + 1$"),
                                                  TeX("$\\cdots$"),
                                                  TeX("$2\\left(\\frac{b(n-1)}{n}\\right)^2 + 1$"),
                                                  TeX("$2b^2 + 1$")), 
                                     y_brake_init = 1, 
                                     y_limit_inf = 0,  
                                     fill_bars = "#d8b365")

approx_below_2x2_1 + approx_above_2x2_1
```

First we need to calculate the area of each rectangle:

$$\frac{b}{n} \left[\mathbf{2}\left[\frac{kb}{n}\right]^2 + \mathbf{1}\right] = \mathbf{2}\frac{k^2b^3}{n^3} + \mathbf{1}\frac{b}{n}$$

Second we need to calculate $s_n$ 

$$\begin{split}
  s_n & = \sum_{k=1}^{n-1} \left[\mathbf{2}\frac{k^2b^3}{n^3} + \mathbf{1}\frac{b}{n}\right] \\
      & = \mathbf{2}\frac{b^3}{n^3} \sum_{k=1}^{n-1} k^2 + \mathbf{1}\frac{b(n-1)}{n} \\
      & = \mathbf{2}\frac{b^3}{n^3} \left[\frac{n^3}{3} - \frac{n^2}{2} + \frac{n}{6} \right] + \mathbf{1}\frac{b(n-1)}{n}
  \end{split}$$
  
Third we need to calculate $S_n$ 

$$\begin{split}
  S_n & = \sum_{k=1}^{n} \left[\mathbf{2}\frac{k^2b^3}{n^3} + \mathbf{1}\frac{b}{n}\right] \\
      & = \mathbf{2}\frac{b^3}{n^3} \sum_{k=1}^{n} k^2 + \mathbf{1}b \\
      & = \mathbf{2}\frac{b^3}{n^3} \left[\frac{n^3}{3} + \frac{n^2}{2} + \frac{n}{6} \right] + \mathbf{1}b
  \end{split}$$
  
Fourth we use the following inequalities where $n \geq 1$

$$\begin{split}
  \frac{n^3}{3} - \frac{n^2}{2} + \frac{n}{6} & < \frac{n^3}{3} < \frac{n^3}{3} + \frac{n^2}{2} + \frac{n}{6} \\
  \mathbf{2}\frac{b^3}{n^3} \left[\frac{n^3}{3} - \frac{n^2}{2} + \frac{n}{6}\right] + \mathbf{1}b & < \mathbf{2}\frac{b^3}{3} + \mathbf{1}b < \mathbf{2}\frac{b^3}{n^3} \left[\frac{n^3}{3} + \frac{n^2}{2} + \frac{n}{6}\right] + \mathbf{1}b \\
  \mathbf{2}\frac{b^3}{n^3} \left[\frac{n^3}{3} - \frac{n^2}{2} + \frac{n}{6}\right] + \mathbf{1}\frac{b(n-1)}{n} & < \mathbf{2}\frac{b^3}{3} + \mathbf{1}b < \mathbf{2}\frac{b^3}{n^3} \left[\frac{n^3}{3} + \frac{n^2}{2} + \frac{n}{6}\right] + \mathbf{1}b
  s_n & < \mathbf{2}\frac{b^3}{3} + \mathbf{1}b < S_n
  \end{split}$$

Fifth we assume that there is some $A$ such that $s_n < A < S_n$ and prove that $A = \mathbf{2}\frac{b^3}{3} + \mathbf{1}b$ using the following inequalities: $\frac{n^3}{3} - \frac{n^2}{2} + \frac{n}{6} < \frac{n^3}{3}$ and $\frac{n^3}{3} < \frac{n^3}{3} + \frac{n^2}{2} + \frac{n}{6}$

$$\begin{split}
  \frac{n^3}{3} - \frac{n^2}{2} + \frac{n}{6} < \frac{n^3}{3} & \iff \frac{n^3}{3} + \frac{n^2}{2} + \frac{n}{6} < \frac{n^3}{3} + n^2 \\
  & \iff \mathbf{2}\frac{b^3}{n^3}\left[\frac{n^3}{3} + \frac{n^2}{2} + \frac{n}{6}\right] + \mathbf{1}b < \mathbf{2}\frac{b^3}{n^3}\left[\frac{n^3}{3} + n^2\right] + \mathbf{1}b \\
  & \iff S_n < \mathbf{2}\left[\frac{b^3}{3} + \frac{b^3}{n}\right] + \mathbf{1}b
  \end{split}$$
  
  $$\begin{split}
  \frac{n^3}{3} < \frac{n^3}{3} + \frac{n^2}{2} + \frac{n}{6} & \iff \frac{n^3}{3} - n^2 < \frac{n^3}{3} - \frac{n^2}{2} + \frac{n}{6} \\
  & \iff \mathbf{2}\frac{b^3}{n^3}\left[\frac{n^3}{3} - n^2\right] + \mathbf{1}b < \mathbf{2}\frac{b^3}{n^3}\left[\frac{n^3}{3} - \frac{n^2}{2} + \frac{n}{6}\right] + \mathbf{1}b \\
  & \iff \mathbf{2}\left[\frac{b^3}{3} - \frac{b^3}{n}\right] + \mathbf{1}b < s_n \\
  \end{split}$$

Therefore we have also that $\mathbf{2}\left[\frac{b^3}{3} - \frac{b^3}{n}\right] + \mathbf{1}b < A < \mathbf{2}\left[\frac{b^3}{3} + \frac{b^3}{n}\right] + \mathbf{1}b$. If $A > \mathbf{2}\frac{b^3}{3} + \mathbf{1}b$ then $A - \mathbf{2}\frac{b^3}{3} - \mathbf{1}b > 0$ so using $A < \mathbf{2}\left[\frac{b^3}{3} + \frac{b^3}{n}\right] + \mathbf{1}b$ we can conclude that $n < \frac{\mathbf{2}b^3}{A - \mathbf{2}\frac{b^3}{3} - \mathbf{1}b}$ but $n \geq 1$ so there are some values of $n$ where $n < \frac{\mathbf{2}b^3}{A - \mathbf{2}\frac{b^3}{3} - \mathbf{1}b}$ is false. If $A < \mathbf{2}\frac{b^3}{3} + \mathbf{1}b$ then $0 < \mathbf{2}\frac{b^3}{3} + \mathbf{1}b - A$ so using $\mathbf{2}\left[\frac{b^3}{3} - \frac{b^3}{n}\right] + \mathbf{1}b < A$ we can also conclude that $n < \frac{\mathbf{2}b^3}{\mathbf{2}\frac{b^3}{3} + \mathbf{1}b - A}$ and again this is false. The only possibility is that $A = \mathbf{2}\frac{b^3}{3} + \mathbf{1}b$ so there is a unique value $A$ such that $s_n < A < S_n$.

**1.e** Do the same if the ordinate at each $x$ is $ax^2 + c$

In this part we are going to assume that $a > 0$ and $c > 0$. If $a = 0$ then it is not possible to calculate the area of the rectangles and if $a < 0$ we would need to define the concept of a negative area. Also if $c = 0$ it is a case we already cover and if $c < 0$ we would need again to define the concept of a negative area.    

```{r echo=FALSE, warning=FALSE, fig.height=6, fig.width=7, fig.align='center'}

parabola_ax2_c <- function(x) {
    return(2*x^2 + 1)
}

approx_below_ax2_c <- plot_integral_approx(fun = parabola_ax2_c, 
                                     approx_type = "below", 
                                     y_labels = c(TeX("$c$"),
                                                  TeX("$a\\left(\\frac{b}{n}\\right)^2 + c$"),
                                                  TeX("$a{1}{4}\\left(\\frac{2b}{n}\\right)^2 + c$"),
                                                  TeX("$\\cdots$"),
                                                  TeX("$a\\left(\\frac{b(n-1)}{n}\\right)^2 + c$"),
                                                  TeX("$ab^2 + c$")), 
                                     y_brake_init = 1, 
                                     y_limit_inf = 0,   
                                     fill_bars = "#a1d76a")

approx_above_ax2_c <- plot_integral_approx(fun = parabola_ax2_c, 
                                     approx_type = "above", 
                                     y_labels = c(TeX("$c$"),
                                                  TeX("$a\\left(\\frac{b}{n}\\right)^2 + c$"),
                                                  TeX("$a{1}{4}\\left(\\frac{2b}{n}\\right)^2 + c$"),
                                                  TeX("$\\cdots$"),
                                                  TeX("$a\\left(\\frac{b(n-1)}{n}\\right)^2 + c$"),
                                                  TeX("$ab^2 + c$")), 
                                     y_brake_init = 1, 
                                     y_limit_inf = 0,  
                                     fill_bars = "#d8b365")

approx_below_ax2_c + approx_above_ax2_c
```

The process is similar as in **1.d** where we only need to change $\mathbf{2}$ with $\mathbf{a}$ and $\mathbf{1}$ with $\mathbf{c}$.

**2** Modify the region in Figure I.3 by assuming that the ordinate at each $x$ is $x^3$ instead of $x^2$. Draw a new figure.

```{r echo=FALSE, warning=FALSE, fig.height=8, fig.width=7, fig.align='center'}

x3 <- function(x) {
    return(x^3)
}

approx_below_x3 <- plot_integral_approx(fun = x3, 
                                     approx_type = "below", 
                                     y_labels = c(TeX("$\\left(\\frac{b}{n}\\right)^3$"),
                                                  TeX("$\\left(\\frac{2b}{n}\\right)^3$"),
                                                  TeX("$\\cdots$"),
                                                  TeX("$\\left(\\frac{b(n-1)}{n}\\right)^3$"),
                                                  TeX("$b^3$")), 
                                     y_brake_init = 2, 
                                     y_limit_inf = 0,   
                                     fill_bars = "#a1d76a")

approx_above_x3 <- plot_integral_approx(fun = x3, 
                                     approx_type = "above", 
                                     y_labels = c(TeX("$\\left(\\frac{b}{n}\\right)^3$"),
                                                  TeX("$\\left(\\frac{2b}{n}\\right)^3$"),
                                                  TeX("$\\cdots$"),
                                                  TeX("$\\left(\\frac{b(n-1)}{n}\\right)^3$"),
                                                  TeX("$b^3$")), 
                                     y_brake_init = 2, 
                                     y_limit_inf = 0,  
                                     fill_bars = "#d8b365")

approx_below_x3 + approx_above_x3
```

**2a** Use a construction similar to that illustrated in Figure I.3 and show that the outer and inner sums $S_n$ and $s_n$ are given by

$$S_n = \frac{b^4}{n^4}\left[1^3 + 2^3 + \cdots + n^3\right] \; , \; s_n = \frac{b^4}{n^4}\left[1^3 + 2^3 + \cdots + (n-1)^3\right]$$

First we need to calculate the area of each rectangle:

$$\frac{b}{n} \left[\frac{kb}{n}\right]^3 = \frac{k^3b^4}{n^4}$$

Second we need to calculate $s_n$ 

$$\begin{split}
  s_n & = \sum_{k=1}^{n-1} \frac{k^3b^4}{n^4} \\
      & = \frac{b^4}{n^4} \sum_{k=1}^{n-1} k^3
  \end{split}$$
  
Third we need to calculate $S_n$ 

$$\begin{split}
  S_n & = \sum_{k=1}^{n} \frac{k^3b^4}{n^4} \\
      & = \frac{b^4}{n^4} \sum_{k=1}^{n} k^3
  \end{split}$$

**2b** Use the inequalities (which can be proved by mathematical induction; see Section I4.2)

$$1^3 + 2^3 + \cdots + (n-1)^3 < \frac{n^4}{4} < 1^3 + 2^3 + \cdots + n^3$$

to show that $s_n < \frac{b^4}{4} < S_n$ for every $n$, and prove that $\frac{b^4}{4}$ is the only number which lies between $s_n$ and $S_n$ for every $n$

First we use the following inequality where $n \geq 1$

$$\begin{split}
  1^3 + 2^3 + \cdots + (n-1)^3 & < \frac{n^4}{4} & < 1^3 + 2^3 + \cdots + n^3 \\
  \frac{b^4}{n^4} \sum_{k=1}^{n-1} k^3 & < \frac{b^4}{4} & < \frac{b^4}{n^4} \sum_{k=1}^{n} k^3 \\
  s_n & < \frac{b^4}{4} & < S_n
  \end{split}$$

Second we assume that there is some $A$ such that $s_n < A < S_n$ and prove that $A = \frac{b^4}{4}$ using the following inequalities: $s_n < \frac{b^4}{4}$ and $\frac{b^4}{4} < S_n$

If $A > \frac{b^4}{4}$ and $s_n < \frac{b^4}{4}$ then $S_n > \frac{b^4}{4}$ and $0 < \frac{b^4}{4} - s_n$. Therefore because $S_n - s_n > \frac{b^4}{4} - s_n$ this implies that $\frac{b^4}{\frac{b^4}{4} - s_n} > n$ but $n \geq 1$ so there are some values of $n$ where $\frac{b^4}{\frac{b^4}{4} - s_n} > n$ is false. 

If $A < \frac{b^4}{4}$ and $\frac{b^4}{4} < S_n$ then $\frac{b^4}{4} > s_n$ and $0 < S_n - \frac{b^4}{4}$. Therefore because $S_n - \frac{b^4}{4} < S_n - s_n$ this implies that $n < \frac{b^4}{S_n - \frac{b^4}{4}}$ but $n \geq 1$ so there are some values of $n$ where $n < \frac{b^4}{S_n - \frac{b^4}{4}}$ is false.

**2c** What numbers takes the place of $\frac{b^4}{4}$ if the ordinate at each $x$ is $ax^3 + c$?

The process to answer this question is similar to the answer in **2b** and **2a** 

First we need to calculate the area of each rectangle:

$$\frac{b}{n} \left[ a\left[\frac{kb}{n}\right]^{\mathbf{3}} + c\right] = a\frac{k^{\mathbf{3}}b^{\mathbf{4}}}{n^{\mathbf{4}}} + c \frac{b}{n}$$

Second we need to calculate $s_n$ 

$$\begin{split}
  s_n & = \sum_{k=1}^{n-1} \left[a\frac{k^{\mathbf{3}}b^{\mathbf{4}}}{n^{\mathbf{4}}} + c \frac{b}{n}\right] \\
      & = a\frac{b^{\mathbf{4}}}{n^{\mathbf{4}}} \sum_{k=1}^{n-1} k^{\mathbf{3}} + c\frac{b(n-1)}{n}
  \end{split}$$
  
Third we need to calculate $S_n$ 

$$\begin{split}
  S_n & = \sum_{k=1}^{n} \left[a\frac{k^{\mathbf{3}}b^{\mathbf{4}}}{n^{\mathbf{4}}} + c \frac{b}{n}\right] \\
      & = a\frac{b^{\mathbf{4}}}{n^{\mathbf{4}}} \sum_{k=1}^{n} k^{\mathbf{3}} + cb
  \end{split}$$

Fourth we use the following inequality where $n \geq 1$

$$\begin{split}
  1^{\mathbf{3}} + 2^{\mathbf{3}} + \cdots + (n-1)^{\mathbf{3}} & < \frac{n^{\mathbf{4}}}{{\mathbf{4}}} < 1^{\mathbf{3}} + 2^{\mathbf{3}} + \cdots + n^{\mathbf{3}} \\
  a\frac{b^{\mathbf{4}}}{n^{\mathbf{4}}} \sum_{k=1}^{n-1} k^{\mathbf{3}} + cb < a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} + cb & < a\frac{b^{\mathbf{4}}}{n^{\mathbf{4}}} \sum_{k=1}^{n} k^{\mathbf{3}} + cb \\
  a\frac{b^{\mathbf{4}}}{n^{\mathbf{4}}} \sum_{k=1}^{n-1} k^{\mathbf{3}} + c\frac{b(n-1)}{n} & < a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} + cb < a\frac{b^{\mathbf{4}}}{n^{\mathbf{4}}} \sum_{k=1}^{n} k^{\mathbf{3}} + cb \\
  s_n & < a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} + cb < S_n
  \end{split}$$

Sixth we assume that there is some $A$ such that $s_n < A < S_n$ and prove that $A = a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} + cb$ using the following inequalities: $s_n < a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} + cb$ and $a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} + cb < S_n$

If $A > a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} + cb$ and $s_n < a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} + cb$ then $S_n > a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} + cb$ and $0 < a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} + cb - s_n$. Therefore because $S_n - s_n > a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} + cb - s_n$ this implies that $\frac{ab^{\mathbf{4}} + cb}{a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} + cb - s_n} > n$ but $n \geq 1$ so there are some values of $n$ where $\frac{ab^{\mathbf{4}} + cb}{a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} + cb - s_n} > n$ is false. 

If $A < a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} + cb$ and $a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} + cb < S_n$ then $a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} + cb > s_n$ and $0 < S_n - a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} - cb$. Therefore because $S_n - a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} - cb < S_n - s_n$ this implies that $n < \frac{ab^4 + cb}{S_n - a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} - cb}$ but $n \geq 1$ so there are some values of $n$ where $n < \frac{ab^{\mathbf{4}} + cb}{S_n - a\frac{b^{\mathbf{4}}}{{\mathbf{4}}} - cb}$ is false.

**3** The inequalities (I.5) and (I.12) are special cases of the more general inequalities

$$1^{\mathbf{m}} + 2^{\mathbf{m}} + \cdots + (n-1)^{\mathbf{m}} < \frac{n^{\mathbf{m + 1}}}{\mathbf{m + 1}} < 1^{\mathbf{m}} + 2^{\mathbf{m}} + \cdots + n^{\mathbf{m}}$$

that are valid for every integer $n \geq 1$ and every integer ${\mathbf{m}} \geq 1$. Asumme the validity of (I.13) and generalize the results of Exercise 2.

We can generalize this result to $ax^{\mathbf{m}} + c$ where the process is similar as in **2.c**. We only need to change $\mathbf{3}$ with $\mathbf{m}$ and $\mathbf{4}$ with $\mathbf{m + 1}$.

## Some Basic Concepts of the Theory of Sets

### Exercises I 2.5 {#exrI2.5}

**1** Use the roster notation to designate the following sets of real numbers.

$$A = \{x | x^2 - 1 = 0 \} = \{-1, 1 \}$$

## A Set of Axioms for the Real-Number System

### Exercises I 3.3 {#exrI3.3}

### Exercises I 3.5 {#exrI3.5}

### Exercises I 3.12 {#exrI3.12}

## Mathematical Induction, Summation Notation, and Related Topics

### Exercises I 4.4 {#exrI4.4}

### Exercises I 4.7 {#exrI4.7}

### Exercises I 4.9 {#exrI4.9}
